---
title: "Read 3 to ICD-10 mapping"
author: "A Warwick and CY Ung"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    toc: false
    code_folding: hide
editor_options:
  chunk_output_type: console
---

```{r global-options, echo=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(reactable)
library(readxl)
library(crosstalk)
library(targets)
library(codemapper)
library(flextable)

tar_load(all_lkps_maps_raw)
read_ctv3_icd10 <- all_lkps_maps_raw$read_ctv3_icd10
```

```{r}
# utility functions
append_read3_icd10_descriptions <- function(df) {
  # get read3 and icd10 descriptions
  read3_df <- df$read_code %>%
    lookup_codes(code_type = "read3", 
                 unrecognised_codes = "warning") %>%
    select(read_code = code,
           description_read3 = description)
  
  icd10_df <- df$icd10_code %>%
    lookup_codes(code_type = "icd10",
                 unrecognised_codes = "warning") %>%
    select(icd10_code = code,
           description_icd10 = description)
  
  # append descriptions
  list(df,
       read3_df,
       icd10_df) %>%
    reduce(full_join) %>%
    select(contains("read"),
           contains("icd10"),
           everything())
}
```

# Key points

-   Dagger/asterisk ICD10 codes have 'D'/'A' appended respectively in the `read_ctv3_icd10` mapping table. These characters need removing to obtain the actual ICD10 codes that appear in the `icd10_lkp` table and UK Biobank data.
-   Inaccurate mapping of read3 to ICD10 codes when codes with `mapping_status` 'R' are mapped without referring to codes with mapping status 'D'.

# Unrecognised ICD10 codes

```{r}
# only the footer note (which appears under `read_code` in the raw `read_ctv3_icd10`
# mapping table) is not present in the `read_ctv3_lkp` lookup table)
unrecognised_read3_codes <- subset(read_ctv3_icd10$read_code,
       !read_ctv3_icd10$read_code %in% all_lkps_maps_raw$read_ctv3_lkp$read_code)

assertthat::are_equal(unrecognised_read3_codes,
                      "ICD-10 codes, terms and text used by permission of WHO, from: International Statistical Classification of Diseases and Related Health Problems, Tenth Revision (ICD-10). Vols 1-3. Geneva, World Health Organization, 1992-2016.")
```

All Read 3 codes in the `read_ctv3_icd10` mapping table are present in the `read_ctv3_lkp` lookup table.

```{r}
# identify unrecognised ICD10 codes
unrecognised_icd10_codes <- subset(read_ctv3_icd10$icd10_code,
       !read_ctv3_icd10$icd10_code %in% all_lkps_maps_raw$icd10_lkp$ALT_CODE) %>% 
  unique()
```

There are `r length(unrecognised_icd10_codes)` ICD10 codes in the `read_ctv3_icd10` mapping table that are *not* present in the `icd10_lkp` table (`ALT_CODE` format).

```{r}
# do these all end in 'A' or 'D'? (i.e. 'asterisk' or 'dagger')
assertthat::are_equal(length(unrecognised_icd10_codes),
                      (read_ctv3_icd10 %>% 
                        filter(str_detect(icd10_code,
                                          pattern = "[A|D]$")) %>% 
                        pull(icd10_code) %>% 
                        unique() %>% 
                        length()))
```

These are all either 'asterisk' or 'dagger' codes, which have been appended with 'A' or 'D' respectively. 

```{r}
# remove 'D'/'A' appendices from dagger/asterisk ICD10 codes
read_ctv3_icd10_icd10_codes_modified <- read_ctv3_icd10 %>% 
  reformat_read_ctv3_icd10() %>% 
  pull(icd10_code) %>% 
  unique()

# see if any 'unrecognised' ICD10 codes remain
unrecognised_icd10_codes <- subset(read_ctv3_icd10_icd10_codes_modified,
       !read_ctv3_icd10_icd10_codes_modified %in% all_lkps_maps_raw$icd10_lkp$ALT_CODE)

assertthat::are_equal(length(unrecognised_icd10_codes),
                      0)
```

After removing 'A' or 'D' from the ICD10 codes in mapping table `read_ctv3_icd10`, all ICD10 codes are also present in lookup table `icd10_lkp`.

# Mapping status and refine flag

## Mapping status

The Read 3 to ICD10 mapping table has a number of columns with additional information:

```{r}
read_ctv3_icd10 %>% 
  head() %>% 
  knitr::kable(caption = "First few rows of the Read 3 to ICD10 mapping table")
```

These are described by UK Biobank [resource 592](https://biobank.ndph.ox.ac.uk/ukb/refer.cgi?id=592). The following tables show how `refine_flag` status varies by `mapping_status`. Mappings labeled with `mapping_status` 'E', 'D' and 'G' have the lowest proportion labelled with `refine_flag` 'M' (i.e. it is mandatory to check mappings labelled with `refine_flag` 'M' against the default)

**Counts:**

```{r}
refine_flag_by_mapping_status <- read_ctv3_icd10 %>% 
  distinct(mapping_status) %>% 
  filter(!is.na(mapping_status)) %>% 
  pull(mapping_status) %>% 
  set_names() %>% 
  map(~ {
    read_ctv3_icd10 %>% 
        filter(mapping_status == .x) %>% 
        group_by(refine_flag) %>% 
      summarise(n_unique_read3 = length(unique(read_code)))
    }) %>% 
  bind_rows(.id = "mapping_status") %>% 
  mutate(refine_flag = paste0("refine_flag_", refine_flag)) %>% 
  pivot_wider(names_from = refine_flag,
              values_from = n_unique_read3) %>% 
  mutate(Total_unique_read3 = rowSums(across(refine_flag_C:refine_flag_P)))

knitr::kable(refine_flag_by_mapping_status)
```

**Percentages:**

```{r}
refine_flag_by_mapping_status_pct <- refine_flag_by_mapping_status %>% 
  mutate(across(refine_flag_C:refine_flag_P,
                ~ round(.x / Total_unique_read3 * 100, digits = 3)))

knitr::kable(refine_flag_by_mapping_status_pct)
```

## Refine flag

Mappings are categorised under the following `refine_flag` labels:

| `refine_flag` | Description                                  |
|---------------|----------------------------------------------|
| `C`           | Completely refined                           |
| `M`           | Mandatory to refine further                  |
| `P`           | Possible but not mandatory to refine further |

| The `refine_flag` denotes whether or not the target code is sufficiently detailed to be acceptable. 3-character ICD codes are usually not acceptable, for example. Covers addition of 4th and 5th digit extensions in ICD, 4th character in OPCS-4.  |
|------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|

```{r}
# get all possible combinations of `mapping_status`, `refine_flag` and
# `add_code_flag`
mapping_metadata_vars <- c("mapping_status",
  "refine_flag",
  "add_code_flag")

mapping_metadata_vars_combinations <- mapping_metadata_vars %>% 
  set_names() %>% 
  
  # get unique values for each variable
  map(~ read_ctv3_icd10 %>% 
        filter(!is.na(.data[[.x]])) %>% 
        pull(.data[[.x]]) %>% 
               unique()) %>% 
  
  # generate all possible combinations
  crossing(!!!.) %>% 
  
  # perform filtering join for only combinations that are actually present in
  # mapping table
  semi_join(read_ctv3_icd10,
            by = mapping_metadata_vars)

# convert this to a named list (where names describe the combination)
mapping_metadata_vars_combinations_list <- mapping_metadata_vars_combinations %>% 
  unite(col = "combination_label",
        everything(),
        remove = FALSE) %>% 
  mutate(combination_label = paste0("mapping_status/refine_flag/add_code_flag: ",
                                    combination_label))

mapping_metadata_vars_combinations_list <-
  split(
    mapping_metadata_vars_combinations_list,
    mapping_metadata_vars_combinations_list$combination_label
  ) %>% 
  map(~ select(.x, -combination_label))

mapping_metadata_vars_combinations_list
```

Here are some example codes for each refine flag

```{r results='asis'}
mapping_metadata_vars_combinations_list %>% 
  map(~ right_join(.x,
                   read_ctv3_icd10,
                   by = mapping_metadata_vars) %>% 
        head(1) %>% 
        append_read3_icd10_descriptions())
```


```{r}
# for `refine_flag` 'M', get one example code mapping per `mapping_status`
refine_flag_m_examples <- read_ctv3_icd10 %>% 
  filter(refine_flag == "M") %>% 
  group_by(mapping_status) %>% 
  slice(1) %>% 
  ungroup()
```

```{r}
# filter for all rows that include these codes, and append descriptions
read_ctv3_icd10 %>% 
  filter(read_code %in% refine_flag_m_examples$read_code |
           icd10_code %in% refine_flag_m_examples$icd10_code) %>% 
  append_read3_icd10_descriptions() %>% 
  reactable(filterable = TRUE,
            searchable = TRUE,
            showPageSizeOptions = TRUE)
```

# One-to-many Read 3-to-ICD10 mappings

```{r}
# identify read3 codes that map to more than one icd10 code
one_to_many_read3_to_icd10_maps <- read_ctv3_icd10 %>% 
  group_by(read_code) %>% 
  mutate(n = n()) %>% 
  ungroup() %>% 
  filter(n > 1)
```

There are `r n_distinct(one_to_many_read3_to_icd10_maps$read_code)` Read 3 codes that map to more than one ICD10 code. The number of these codes per `mapping_status` is summarised below. Note that none of these have `mapping_status` `E`:

```{r}
one_to_many_read3_to_icd10_maps %>% 
  group_by(mapping_status) %>% 
  summarise(n_unique_read3 = n_distinct(read_code)) %>% 
  knitr::kable()
```

The number of these codes per `refine_flag` is summarised below:

```{r}
one_to_many_read3_to_icd10_maps %>% 
  group_by(refine_flag) %>% 
  summarise(n_unique_read3 = n_distinct(read_code)) %>% 
  knitr::kable()
```

Here are some example codes with their descriptions. Note that for each Read 3 code there is one marked as `D` (i.e. the 'default' mapping):

```{r}
# see some examples
one_to_many_read3_to_icd10_maps_examples <- one_to_many_read3_to_icd10_maps %>% 
  head(10)

one_to_many_read3_to_icd10_maps %>% 
  filter(read_code %in% one_to_many_read3_to_icd10_maps_examples$read_code) %>% 
  append_read3_icd10_descriptions() %>% 
  reactable(filterable = TRUE,
            showPageSizeOptions = TRUE)
```

```{r}
# get some example codes with more than one mapping labelled 'D'
one_to_many_read3_to_icd10_maps_2_more_D <- one_to_many_read3_to_icd10_maps %>% 
  group_by(read_code) %>% 
  summarise(n_mapping_status_D = sum(mapping_status == "D")) %>% 
  filter(n_mapping_status_D > 1) %>% 
  head() %>% 
  pull(read_code)

# see descriptions for these codes
read_ctv3_icd10 %>% 
  filter(read_code %in% one_to_many_read3_to_icd10_maps_2_more_D) %>% 
  append_read3_icd10_descriptions() %>% 
  reactable(filterable = TRUE,
            showPageSizeOptions = TRUE)
```

# Modified ICD10 codes

-   Get unrecognised ICD10 codes

```{r}
read_ctv3_icd10 %>%
  mutate(icd10_code_lkp =
           str_remove(icd10_code, pattern = paste0(str_c(
             LETTERS[-24], sep = "", collapse = "|"
           ), "$")))
```

# TODO

The read3 code `XaIP9` mistakenly maps 'sebaceous cyst' to various ICD10 codes that code for 'Other specified disorders of eyelid' (`H028`), 'Other specified disorders of male genital organs' (`N508`), 'Other benign mammary dysplasias' (`N608`) and 'Other specified conditions associated with female genital organs...' (`N948`)

These inaccurate codes are coded 'R' under 'mapping_status' in 'read_ctv3_icd10'.

This field naming system is also used in read_ctv3_icd9.

```{r}
mapped_codes <-map_codes(codes = "XaIP9",
          from = "read3",
          to = "icd10",
          all_lkps_maps = "all_lkps_maps.db",
          standardise_output = FALSE)

icd10_descriptions <-map_codes(codes = "XaIP9",
          from = "read3",
          to = "icd10",
          all_lkps_maps = "all_lkps_maps.db",
          standardise_output = TRUE) %>% dplyr::rename("icd10_code" = "code",
                                                       "icd10_description" = "description")

joined <- dplyr::left_join(mapped_codes, icd10_descriptions)

read3_descriptions<-codemapper::lookup_codes(joined$read_code, code_type = "read3") 

result <- joined %>% dplyr::mutate(read_description = read3_descriptions$description) %>% dplyr::select(read_code,
                                                                         read_description,
                                                                         icd10_code,
                                                                         icd10_description)
 
```

# Transformed table
