---
title: "Reformatting UKB 'all_lkps_maps.xlsx'"
author: "Alasdair Warwick"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

```{r global-options, echo=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=FALSE)
```

```{r}
library(tidyverse)
library(reactable)
library(readxl)
library(crosstalk)
library(targets)

devtools::load_all()

config <- configr::read.config("config.ini")

all_lkps_maps_raw <- tar_read(all_lkps_maps_raw)
all_lkps_maps <- tar_read(all_lkps_maps)
bnf_dmd <- tar_read(bnf_dmd)
```

# Remove irrelevant rows 

## Last few rows of each sheet {.tabset .tabset-pills}

2 columns shown only

```{r results='asis'}
iwalk(all_lkps_maps_raw,
     ~ {
       cat("### ", .y, "\n\n")
       
       print(knitr::kable(tail(.x[, 1:2])))
       
       cat("\n\n")
     })
```

## Cannot just remove `NA` in column 2

```{r}
# remove irrelavant rows
all_lkps_maps_rm_na_col2 <- all_lkps_maps_raw %>% 
  map(~ .x[!is.na(.x[[2]]), ])
```

```{r}
# compare nrows before and after this step
all_lkps_maps_rm_na_col2_nrows <- all_lkps_maps_rm_na_col2 %>% 
  map(nrow) %>% 
  as.data.frame() %>% 
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column("Sheet") %>% 
  rename(`N rows rm NAs in col 2` = V1)

all_lkps_maps_nrows <- all_lkps_maps_raw %>% 
  map(nrow) %>% 
  as.data.frame() %>% 
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column("Sheet") %>% 
  rename(`N rows raw` = V1)

# combine and compare
all_lkps_maps_nrows %>% 
  full_join(all_lkps_maps_rm_na_col2_nrows,
            by = "Sheet") %>% 
  mutate(Difference = `N rows raw` - `N rows rm NAs in col 2`) %>% 
  knitr::kable()
```

## Instead, determine N rows to remove manually

```{r}
# remove last few rows from each sheet (n to remove determined manually)
all_lkps_maps_reformatted <- remove_irrelevant_rows_all_lkps_maps(all_lkps_maps_raw)

# compare row numbers
all_lkps_maps_reformatted_nrows <- all_lkps_maps_reformatted %>% 
  map(nrow) %>% 
  as.data.frame() %>% 
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column("Sheet") %>% 
  rename(`N rows reformatted` = V1)

all_lkps_maps_nrows %>% 
  full_join(all_lkps_maps_reformatted_nrows,
            by = "Sheet") %>% 
  mutate(Difference = `N rows raw` - `N rows reformatted`) %>% 
  knitr::kable()
```

# `bnf_lkp`

BNF codes in the UK Biobank linked primary care dataset are of variable length (see [UKB resource 591](https://biobank.ndph.ox.ac.uk/ukb/refer.cgi?id=591)).

`extend_bnf_lkp()` creates all possible versions for BNF codes in `bnf_lkp`. This facilitates selecting all drugs under a certain chapter/section/etc e.g. to get all BNF codes for all medications in the 'Treatment Of Glaucoma' section (i.e. BNF codes starting with `1106`). This table shows the results for the first drug in this section:

```{r}
all_lkps_maps_raw$bnf_lkp %>% 
  filter(BNF_Section == "Treatment Of Glaucoma") %>% 
  extend_bnf_lkp() %>% 
  head(n = 8) %>% 
  reactable(filterable = TRUE,
            resizable = TRUE,
            defaultPageSize = 8)
```

# `bnf_read2`

All BNF codes in the `bnf_code` col are 8 digits long. Digit 7 corresponds to BNF sub paragraph - this is expanded to 2 digits by prefixing with a `0` in `bnf_code`. 

Note that there are a number of Read 2 codes that have not been mapped to BNF (`NA` in the table below):

```{r}
all_lkps_maps$read_v2_drugs_bnf %>%
  mutate("Digits 7 and 8, bnf_code col" = str_sub(
    str_remove_all(bnf_code, pattern = "\\."),
    start = 7,
    end = 8
  )) %>%
  count(`Digits 7 and 8, bnf_code col`) %>%
  knitr::kable()
```

...in most cases, `BNF_Paragraph` = `BNF_Subparagraph`. For an example of where these can differ, see BNF paragraph `Renin-Angiotensin System Drugs`.

`extend_read_v2_drugs_bnf()` appends Read 2 and BNF descriptions to the `read_v2_drugs_bnf` table, the first few rows are shown below:

```{r}
extend_read_v2_drugs_bnf(all_lkps_maps) %>% 
  filter() %>% 
  reactable(filterable = TRUE,
            resizable = TRUE,
            defaultPageSize = 8)
```

# `bnf_dmd`

This is downloaded from the [NHSBSA website](https://www.nhsbsa.nhs.uk/prescription-data/understanding-our-data/bnf-snomed-mapping). There are no 'extra' rows that need removing at the bottom.

To map BNF codes to DM+D:

```{r}
# get BNF codes for glaucoma meds
glaucoma_bnf_codes <- all_lkps_maps_raw$bnf_lkp %>% 
  filter(BNF_Section == "Treatment Of Glaucoma") %>% 
  pull(BNF_Presentation_Code)

# are all these codes present in `bnf_dmd`?
glaucoma_bnf_codes_not_in_bnf_dmd <- subset(glaucoma_bnf_codes,!glaucoma_bnf_codes %in% bnf_dmd$bnf_codes)
```

```{r results='asis'}
if (!rlang::is_empty(glaucoma_bnf_codes_not_in_bnf_dmd)) {
  cat(paste0("> **Note:** There are ",
             length(glaucoma_bnf_codes_not_in_bnf_dmd),
             " out of ",
             length(glaucoma_bnf_codes),
             " glaucoma medication codes that are not present in `bnf_dmd`"))
}
```

```{r}
# map to DM+D
bnf_dmd %>% 
  filter(bnf_code %in% glaucoma_bnf_codes) %>% 
  head()
```

